---
title: "DXT-c_pret"
author: "Kristen Johnson"
date: "2025-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# FOR TRANSFORMING DXT SPEECH COLUMN DATA TO C_PRET COLUMN DATA
We don't need all the decontextualized speech codes, just the ones on pretend play. the c_pret column is what is found in other pretend play coding, so I've adopted the same nomenclature.

```{r}
# Load necessary libraries
library(readxl)   # For reading Excel files
library(dplyr)    # For binding and data manipulation
library(lubridate)  # For handling datetime columns (if needed)
```

Read in the P3_H8 CSV file
```{r}

p3_h8 <- read.csv("/PN_Datasets/Pretend_Play/CSVs_of_Combined_Data_PP/P3_H8_cwpu.csv")

```

Add the c_pret column
```{r}
p3_h8 <- p3_h8 %>%
  mutate(c_pret = ifelse(grepl("p", Child.DXT, ignore.case = TRUE), 1, 0)) 

```

Write a new .csv for P3_H8
```{r}
write.csv(p3_h8, "/PN_Datasets/Pretend_Play/CSVs_of_Combined_Data_PP/P3_H8_ALL.csv", row.names = FALSE)
```

Read in the P2_H8 CSV file
```{r}
p2_h8 <- read.csv("PN_Datasets/Pretend_Play/CSVs_of_Combined_Data_PP/P2_H8_cwpu.csv")
```

Add the c_pret column
```{r}
p2_h8 <- p2_h8 %>%
  mutate(c_pret = ifelse(grepl("p", Child.DXT, ignore.case = TRUE), 1, 0)) 
```

write a new .csv for P2_H8
```{r}
write.csv(p2_h8, "PN_Datasets/Pretend_Play/CSVs_of_Combined_Data_PP/P2_H8_ALL.csv", row.names = FALSE)
```

