---
title: "P3_H8_pp_cform"
author: "Kristen Johnson"
date: "2025-02-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# NUMBER & DISTIBUTION OF GESTURES W/I PRETEND PLAY INSTANCES FOR P3_H8 KIDDOS
I needed to see if there were enough gestures within the P3 group (the population less likely to use gestures and/or exhibit instances of pretend play) to justify moving forward with my original study questions.

## read in csv
```{r}
P3_H8_ALL <- read_csv('/Users/kristenjohnson/KristenWorkingDirectory/Narrative Data/CSVs_of_Combined_Data/P3_H8_ALL.csv')
```

## filter for columns on pretend play (c_pret) and gesture (c_form) & count overlaps
```{r}

# Filter for rows where c_pret is 1, then count occurrences of c_form
PP_gest_overlap <- P3_H8_ALL %>%
  filter(c_pret == 1 & !is.na(c_form)) %>%  # Only keep rows where col1 is 5 and c_utt has any value
  count(c_form)  # Count occurrences of c_form

# View the result
print(PP_gest_overlap)

```

## get summary statistics for the number of overlaps
```{r}
library(dplyr)

# Generate summary statistics for the count column 'n'
PP_gest_overlap %>%
  summarise(
    mean_count = mean(n, na.rm = TRUE),
    median_count = median(n, na.rm = TRUE),
    sd_count = sd(n, na.rm = TRUE),
    min_count = min(n, na.rm = TRUE),
    max_count = max(n, na.rm = TRUE),
    total_count = sum(n, na.rm = TRUE),
    num_unique_forms = n()
  )

```

## histogram of average frequency of overlaps per child
```{r}
library(ggplot2)

# Makes a histogram to visualize how frequently different counts of gesture/pretend play overlap appear.
ggplot(PP_gest_overlap, aes(x = n)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Gesture During Play Counts",
       x = "Count of Gesture During Play",
       y = "Frequency") +
  theme_minimal()

```

## boxplot of overlaps
```{r}

# A boxplot shows the median, quartiles, and potential outliers.
ggplot(PP_gest_overlap, aes(y = n)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Boxplot of Gesture During Play Counts",
       y = "Gesture During Play") +
  theme_minimal()

```

## glimpse at different types of gesture used
```{r}

# If you want to see the count of each c_form, use a bar chart.
ggplot(PP_gest_overlap, aes(x = reorder(c_form, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkorange", color = "black") +
  labs(title = "Counts of Gesture During Play Occurrences",
       x = "gesture",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels if needed

```


