---
title: "New_Narrative_Data_Wrangling"
author: "Kristen Johnson"
date: "2025-04-24"
output: html_document
---


```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(stringr)
library(tidyverse)
library(here)
```
# Narrative scores and codes:
1. No structure (**NS**)- only one clause with or without a verb. 
2. Descriptive story (**DS**)- a series of clauses that do not report actions constrained by temporality.
3. Action story (**AC**)-  includes actions *with a temporal order*, but story events are not causally organized.
4. Reactive story (**RS**)- includes actions that are causally organized.
5. Incomplete goal-based story (**IGB**)- contains a goal statement, but no attempt or outcome following the goal.
6. Incomplete goal-based with attempt (**IGBA**)- contains a goal statement, and attempt(s), but not outcome following the goal.
7. Complete goal-based story with one episode implicit goal (**GB1-IG**)- includes temporal and causal structure, as well as *implicitly*-stated goal of the protagonist, *one* series of attempts to achieve the goal, and an outcome of these attempts.
8. Complete goal-based story with one episode explicit goal (**GB1-EG**)- includes temporal and causal structure, as well as *explicitly*-stated goal of the protagonist, *one* series of attempts to achieve the goal, and an outcome of these attempts.
9. Complete goal-based story with multiple episodes implicit goal (**GBM-IG**)- includes *multiple* goal-attempt-outcome sequences and frequently more than one *implicitly*-stated goal. 
10. Complete goal-based story with multiple episodes explicit goal (**GBM-EG**)- includes *multiple* goal-attempt-outcome sequences. with at least one of the goals stated *explicitly*.

# Load in data & convert narrative scores to numbers:
```{r}
# Read in the data
Narrative_Results_Simple <- read_xlsx(here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Simple.xlsx"))

# Get column names of the data frame
column_names <- colnames(Narrative_Results_Simple)

# Print the column names
print(column_names)

# Define the abbreviation-to-number mapping
replacements <- c("NS" = 1, "DS" = 2, "AC" = 3, "RS" = 4, "IGB" = 5, "IGBA" = 6, "GB1-IG" = 7, "GB1-EG" = 8, "GBM-IG" = 9, "GBM-EG" = 10)

# Replace abbreviations in multiple columns and handle NA
Narrative_Results_Numeric <- Narrative_Results_Simple %>%
  mutate(across(c("FoxHOS07", "AliceHOS07", "AlanHOS07",  "FoxHOS08",  "AlanHOS08", "FoxHOS09", "AlanHOS09" ), ~ case_when(
    . == "NS" ~ 1,
    . == "DS" ~ 2,
    . == "AC" ~ 3,
    . == "RS" ~ 4,
    . == "IGB" ~ 5,
    . == "IGBA" ~ 6,
    . == "GB1-IG" ~ 7,
    . == "GB1-EG" ~ 8,
    . == "GBM-IG" ~ 9,
    . == "GBM-EG" ~ 10,
    TRUE ~ as.numeric(NA)  # In case of any unmatched value, set to NA
  ))) %>%
  mutate(across(c("FoxHOS07", "AliceHOS07", "AlanHOS07",  "FoxHOS08",  "AlanHOS08", "FoxHOS09", "AlanHOS09"), ~ replace_na(., 0)))  # Replace NA with 0 in the selected columns

# Write the updated data frame into a csv
write_csv(Narrative_Results_Numeric, here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Numeric.csv"))
```


```{r}
Narrative_Results_Final <- Narrative_Results_Numeric %>%
  mutate(
    avg07 = rowMeans(select(., ends_with("07")), na.rm = TRUE),
    avg08 = rowMeans(select(., ends_with("08")), na.rm = TRUE),
    avg09 = rowMeans(select(., ends_with("09")), na.rm = TRUE),
    max_avg = pmax(avg07, avg08, avg09, na.rm = TRUE)
  ) %>%
  rename(participant_id = ChildID)

print(Narrative_Results_Final)

```


```{r}
write_csv(Narrative_Results_Final, here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Final.csv"))
```

