---
title: "PretendPlay_Gesture_Analysis_ALLyears"
author: "Kristen Johnson"
date: "2025-04-25"
output:
  pdf_document: default
  html_document: default
---

# SETUP
## load packages
```{r, echo=FALSE}
# Install required packages if you don't have them
# Remove the comment symbols (#) to run these lines
# install.packages("tidyverse")
 #install.packages("car")
 #install.packages("effectsize")
 #install.packages("rstatix")
#install.packages("e1071")
# Load required package if needed
# install.packages("rstatix")

# Load necessary packages
library(tidyverse)
library(car)       # For Levene's test and Type III SS
library(effectsize) # For effect sizes
library(rstatix)    # For convenient statistical functions
library(here)
library(extrafont)  # Ensures Times New Roman is available
library(e1071)
library(rstatix)
```

## Read in each dataset and add Year column
```{r}
child_summary_H10 <- read_csv(here("PN_Datasets", "Pretend_Play", "CSVs_of_Combined_Data_PP", "child_summary_H10.csv"))
child_summary_H10$Year <- 10

child_summary_H8 <- read_csv(here("PN_Datasets", "Pretend_Play", "CSVs_of_Combined_Data_PP", "child_summary_H8.csv"))
child_summary_H8$Year <- 8

child_summary_H7 <- read_csv(here("PN_Datasets", "Pretend_Play", "CSVs_of_Combined_Data_PP", "child_summary_H7.csv"))
child_summary_H7$Year <- 7
```

## Combine data across ALL THREE YEARS
```{r}
# Combine all datasets into one data frame
pp_gesture_data_ALLyears <- bind_rows(child_summary_H10, child_summary_H8, child_summary_H7)

# View the combined dataset
head(pp_gesture_data_ALLyears)
```

## Check for Skewness
```{r}
 #Calculate skewness
skewness(pp_gesture_data_ALLyears$rep_gestures, na.rm = TRUE)

# Histogram of rep_gestures
ggplot(pp_gesture_data_ALLyears, aes(x = rep_gestures)) +
  geom_histogram(binwidth = 1, fill = "gray80", color = "black") +
  labs(
    title = "Distribution of Representational Gesture Counts",
    x = "Number of Representational Gestures",
    y = "Count of Participants"
  ) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.border = element_rect(color = "black", fill = NA),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

```
# Correlations in Skewed Data
## Mann-Whitney U test on Rep Gestures (instead of ANOVA, bc highly skewed data)
```{r}
# Step 1: Run the Mann-Whitney U test properly
mw_result <- wilcox.test(rep_gestures ~ groupstatus, 
                         data = pp_gesture_data_ALLyears, 
                         exact = FALSE)

# View the output if you want
print(mw_result)

# Step 2: Calculate total N
n1 <- pp_gesture_data_ALLyears %>%
  filter(groupstatus == "TD") %>%
  nrow()

n2 <- pp_gesture_data_ALLyears %>%
  filter(groupstatus == "PL") %>%
  nrow()

N <- n1 + n2

# Step 3: Approximate Z from p-value
# p.value comes from mw_result
z_approx <- qnorm(mw_result$p.value / 2, lower.tail = FALSE)

# Step 4: Calculate r
r_effect_size <- z_approx / sqrt(N)

# Step 5: Print result
cat("Mann-Whitney U result: W =", mw_result$statistic, 
    ", p =", signif(mw_result$p.value, 3), 
    ", r =", round(r_effect_size, 2), "\n")

# get medians for reporting (due to skewness)
pp_gesture_data_ALLyears %>%
  group_by(groupstatus) %>%
  summarise(Median_RepGest = median(rep_gestures, na.rm = TRUE))
```

## Mann-Whitney U test on Total Gestures
```{r}
# Step 1: Run the Mann-Whitney U test properly
mw_result <- wilcox.test(total_gestures ~ groupstatus, 
                         data = pp_gesture_data_ALLyears, 
                         exact = FALSE)

# View the output if you want
print(mw_result)

# Step 2: Calculate total N
n1 <- pp_gesture_data_ALLyears %>%
  filter(groupstatus == "TD") %>%
  nrow()

n2 <- pp_gesture_data_ALLyears %>%
  filter(groupstatus == "PL") %>%
  nrow()

N <- n1 + n2

# Step 3: Approximate Z from p-value
# p.value comes from mw_result
z_approx <- qnorm(mw_result$p.value / 2, lower.tail = FALSE)

# Step 4: Calculate r
r_effect_size <- z_approx / sqrt(N)

# Step 5: Print result
cat("Mann-Whitney U result: W =", mw_result$statistic, 
    ", p =", signif(mw_result$p.value, 3), 
    ", r =", round(r_effect_size, 2), "\n")

# get medians for reporting (due to skewness)
pp_gesture_data_ALLyears %>%
  group_by(groupstatus) %>%
  summarise(Median_TotGest = median(total_gestures, na.rm = TRUE))
```

## Mann-Whitney U test on Proportion of Rep to Total Gesture
```{r}
# Mann-Whitney U for proportion of representative gestures
wilcox.test(prop_rep ~ groupstatus, 
            data = pp_gesture_data_ALLyears, 
            exact = FALSE)

# Calculate N
n1 <- pp_gesture_data_ALLyears %>% filter(groupstatus == "TD") %>% nrow()
n2 <- pp_gesture_data_ALLyears %>% filter(groupstatus == "PL") %>% nrow()
N <- n1 + n2

# Run Mann-Whitney test
mw_result_prop <- wilcox.test(prop_rep ~ groupstatus, 
                              data = pp_gesture_data_ALLyears, 
                              exact = FALSE)

# Approximate Z from p-value
z_approx_prop <- qnorm(mw_result_prop$p.value / 2, lower.tail = FALSE)

# Calculate r
r_effect_size_prop <- z_approx_prop / sqrt(N)

# Print results
cat("Mann-Whitney U result for proportion: W =", mw_result_prop$statistic, 
    ", p =", signif(mw_result_prop$p.value, 3), 
    ", r =", round(r_effect_size_prop, 2), "\n")

# get medians for reporting (due to skewness)
pp_gesture_data_ALLyears %>%
  group_by(groupstatus) %>%
  summarise(Median_PropRep = median(prop_rep, na.rm = TRUE))
```

## APA Table of ALL Results
```{r}
# Load packages
library(dplyr)
library(gt)

# Create tidy data frame
gesture_results <- tibble::tibble(
  Variable = c("Total Gestures", "Representative Gestures", "Proportion of Representative Gestures"),
  PL_Median = c(3, 0, 0.00),
  TD_Median = c(6, 1, 0.08),
  W = c(573, 508, 277.5),
  p = c(0.235, 0.036, 0.011),
  r = c(0.13, 0.23, 0.29)
) %>%
  mutate(
    sig = ifelse(p < 0.05, TRUE, FALSE),
    p_formatted = ifelse(p < .001, "< .001", sprintf("%.3f", p)),
    r_formatted = sprintf("%.2f", r)
  )

# Build APA-style table: only bold p-values
gesture_results %>%
  select(Variable, PL_Median, TD_Median, W, p_formatted, r_formatted, sig) %>%
  rename(
    `Variable` = Variable,
    `Median (PL)` = PL_Median,
    `Median (TD)` = TD_Median,
    `W` = W,
    `p` = p_formatted,
    `r` = r_formatted
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Table X**"),
    subtitle = md("_Mann-Whitney U Tests Comparing Gesture Measures Between Groups_")
  ) %>%
  tab_style(
    style = cell_text(font = "Times New Roman"),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = p,  # Bold only p-values
      rows = sig == TRUE
    )
  ) %>%
  tab_source_note(
    source_note = md("**Note.** _W_ = Wilcoxon rank sum test statistic. _r_ = effect size (rank biserial correlation). Bolded p-values indicate p < .05.")
  ) %>%
  fmt_number(
    columns = c(`Median (PL)`, `Median (TD)`, `W`, `r`),
    decimals = 2
  ) %>%
  cols_align(
    align = "left", columns = "Variable"
  ) %>%
  cols_align(
    align = "center", columns = c(`Median (PL)`, `Median (TD)`, `W`, `p`, `r`)
  ) %>%
  tab_options(
    table.font.size = "small",
    data_row.padding = px(2),
    table.align = "center",
    table.font.names = "Times New Roman"
  ) %>%
  cols_hide(columns = c(sig))  # Hide significance helper column

```

## Convert gt to flextable
```{r}
# Load required libraries
library(dplyr)
library(flextable)
library(officer)

# Step 1: Prepare your formatted data (similar to what you used in gt)
gesture_results_flex <- gesture_results %>%
  select(Variable, PL_Median, TD_Median, W, p_formatted, r_formatted, sig) %>%
  rename(
    `Variable` = Variable,
    `Median (PL)` = PL_Median,
    `Median (TD)` = TD_Median,
    `W` = W,
    `p` = p_formatted,
    `r` = r_formatted
  )

# Step 2: Create flextable
ft <- flextable(gesture_results_flex) %>%
  bold(i = ~ sig == TRUE, j = "p", bold = TRUE) %>%   # Bold p-values < .05
  set_caption("Table X. Mann-Whitney U Tests Comparing Gesture Measures Between Groups") %>%
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  align(j = 2:6, align = "center", part = "all") %>%
  align(j = 1, align = "left", part = "all") %>%
  padding(padding = 2, part = "all")

# Step 3: Export to Word
doc <- read_docx() %>%
  body_add_par("Results Table", style = "heading 1") %>%
  body_add_flextable(ft) %>%
  body_add_par("Note. W = Wilcoxon rank sum test statistic. r = effect size (rank biserial correlation). Bolded p-values indicate p < .05.", style = "Normal")

print(doc, target = "Gesture_Results_Table.docx")

browseURL("Gesture_Results_Table.docx")


```


## Means & SD's for ALL
```{r}
# Load dplyr
library(dplyr)

# Assuming your gesture_results or full dataset is called pp_gesture_data_ALLyears or similar
# Here’s the simple code:

# Calculate means and SDs by group
gesture_summary_stats <- pp_gesture_data_ALLyears %>%
  group_by(groupstatus) %>%
  summarize(
    Mean_Total_Gestures = mean(total_gestures, na.rm = TRUE),
    SD_Total_Gestures = sd(total_gestures, na.rm = TRUE),
    Mean_Rep_Gestures = mean(rep_gestures, na.rm = TRUE),
    SD_Rep_Gestures = sd(rep_gestures, na.rm = TRUE),
    Mean_Prop_Rep = mean(prop_rep, na.rm = TRUE),
    SD_Prop_Rep = sd(prop_rep, na.rm = TRUE),
    n = n()
  )

# View the summary
gesture_summary_stats

```

## Violin Plot of Representative Gestures X Group Status
```{r}
# Load libraries
library(ggplot2)
library(dplyr)

# Simple violin + jitter plot of representative gestures by group
pp_gesture_data_ALLyears %>%
  ggplot(aes(x = groupstatus, y = rep_gestures, fill = groupstatus)) +
  geom_violin(trim = FALSE, alpha = 0.6) +    # Violin for distribution
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +  # Individual points
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white", color = "black") +  # Mean as white diamond
  labs(
    title = "Figure 1",
    subtitle = "Distribution of Representative Gestures During Pretend Play",
    x = "Group",
    y = "Number of Representative Gestures"
  ) +
  scale_fill_grey(start = 0.4, end = 0.8) +  # Subtle grayscale if you want APA look
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    text = element_text(family = "Times New Roman"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

## improved violin plot as png
```{r}
# Violin plot of representative gestures by group
p_violin <- pp_gesture_data_ALLyears %>%
  ggplot(aes(x = groupstatus, y = rep_gestures, fill = groupstatus)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3,
               fill = "white", color = "black") +
  labs(
    title = "Figure 1",
    subtitle = "Distribution of Representative Gestures During Pretend Play by Group",
    x = "Group",
    y = "Number of Representative Gestures"
  ) +
  scale_fill_grey(start = 0.4, end = 0.8) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

# Save as PNG
ggsave("FigureX_ViolinPlot_RepGestures.png", plot = p_violin,
       width = 7, height = 6, dpi = 300, units = "in", bg = "white")

#Figure 1. Distribution of representative gestures produced during pretend play by group. Violin plots show the full distribution of values, with individual participant points overlaid. White diamonds indicate group means. Children with perinatal lesions (PL) showed lower central tendency and greater clustering at zero relative to typically developing (TD) peers.
```


## Box Plot of Representative Gestures X Group Status
```{r}
# APA-style boxplot of representative gestures by group
pp_gesture_data_ALLyears %>%
  ggplot(aes(x = groupstatus, y = rep_gestures, fill = groupstatus)) +
  geom_boxplot(width = 0.5, outlier.shape = NA, alpha = 0.6) +  # No outlier points (APA-style boxplots often omit)
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +  # Add individual points
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white", color = "black") +  # Mean as white diamond
  labs(
    title = "Number of Representative Gestures During Pretend Play",
    x = "Group",
    y = "Representative Gestures"
  ) +
  scale_fill_grey(start = 0.4, end = 0.8) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    text = element_text(family = "Times New Roman"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

## Bar Graph of Means for Poportion of Rep Gest
```{r}
# Create clean bar graph of proportion of representative gestures
pp_gesture_data_ALLyears %>%
  ggplot(aes(x = groupstatus, y = prop_rep, fill = groupstatus)) +
  stat_summary(fun = mean, geom = "bar", width = 0.6, color = "black", alpha = 0.7) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(
    title = NULL,
    x = "Group",
    y = "Proportion of Representative Gestures"
  ) +
  scale_fill_grey(start = 0.4, end = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +  # Display y-axis as percentages
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    text = element_text(family = "Times New Roman"),
    axis.title.y = element_text(margin = margin(r = 10)),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Figure X
# Mean proportion of representative gestures during pretend play by group. Error bars represent ±1 standard error of the mean. Although group medians were used for nonparametric inferential testing (Mann-Whitney U), means are presented here for ease of interpretation.

```

## fancier version
```{r}
# Load libraries
library(ggplot2)
library(dplyr)

# Bar graph of proportion of representational gestures
pp_gesture_data_ALLyears %>%
  ggplot(aes(x = groupstatus, y = prop_rep, fill = groupstatus)) +
  stat_summary(fun = mean, geom = "bar", width = 0.6, color = "black", alpha = 0.7) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(
    title = "Mean Proportion of Representational Gestures During Pretend Play by Group",
    
    x = "Group",
    y = "Proportion of Representational Gestures"
  ) +
  scale_fill_grey(start = 0.4, end = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5)
  )

#Figure X. Mean proportion of representational gestures during pretend play by group. Error bars indicate ±1 standard error of the mean. Medians were used for statistical testing due to skewness, but means are displayed here for interpretability.

# Save the most recent plot as a high-resolution PNG
ggsave(
  filename = "FigureX_ProportionRepGesture.png",
  width = 7,
  height = 5,
  dpi = 300,
  units = "in",
  device = "png"
)

```

## second try at fancier version
```{r}
# Assign your plot to an object first
p_prop <- ggplot(pp_gesture_data_ALLyears, aes(x = groupstatus, y = prop_rep, fill = groupstatus)) +
  stat_summary(fun = mean, geom = "bar", width = 0.6, color = "black", alpha = 0.7) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(
    title = "Figure 2",
    subtitle = "Mean Proportion of Representational Gestures During Pretend Play by Group",
    x = "Group",
    y = "Proportion of Representative Gestures"
  ) +
  scale_fill_grey(start = 0.4, end = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5),
    plot.background = element_rect(fill = "white", color = NA), # fix gray background
    panel.background = element_rect(fill = "white", color = NA) # fix panel background
  )

# Save with extra height to avoid cutting off the subtitle
ggsave("Figure3_ProportionRepGesture.png", plot = p_prop, width = 7, height = 6, dpi = 300, units = "in", bg = "white")

```


# OLD CODE
## table opton 1
```{r, eval=FALSE}
# Load required libraries
library(dplyr)
library(gt)

# Create a summary table manually (based on your results so far)
mann_whitney_summary <- tibble::tibble(
  Outcome = c(
    "Total Gestures",
    "Number of Representational Gestures",
    "Proportion of Representational Gestures"
  ),
  W = c(802, 867, 615.5),
  p = c(0.235, 0.036, 0.011),
  r = c(0.13, 0.23, 0.29)
)

# Clean up for APA formatting
mann_whitney_summary <- mann_whitney_summary %>%
  mutate(
    p = ifelse(p < .001, "< .001", sprintf("%.3f", p)),
    r = sprintf("%.2f", r)
  )

# Create the APA-style table
mann_whitney_summary %>%
  gt() %>%
  tab_header(
    title = md("**Table X**"),
    subtitle = md("**Mann-Whitney U Test Results for Gesture Outcomes**")
  ) %>%
  cols_label(
    Outcome = "Outcome Variable",
    W = "W",
    p = "p",
    r = "Effect Size (r)"
  ) %>%
  tab_options(
    table.font.size = "small",
    table.align = "center",
    data_row.padding = px(2)
  )
# Note. W = Wilcoxon rank-sum statistic (equivalent to Mann-Whitney U); r = effect size
```

## table opton 2
```{r, eval=FALSE}
# Load libraries
library(dplyr)
library(gt)

# Step 1: Create summary table
mann_whitney_summary <- tibble::tibble(
  Outcome = c(
    "Total Gestures",
    "Number of Representational Gestures",
    "Proportion of Representational Gestures"
  ),
  W = c(802, 867, 615.5),
  p_numeric = c(0.235, 0.036, 0.011),
  r = c(0.13, 0.23, 0.29)
)

# Step 2: Format nicely for APA
mann_whitney_summary <- mann_whitney_summary %>%
  mutate(
    p = ifelse(p_numeric < .001, "< .001", sprintf("%.3f", p_numeric)),
    r = sprintf("%.2f", r)
  )

# Step 3: Build APA table with bolded significant rows
mann_whitney_summary %>%
  select(Outcome, W, p, r, p_numeric) %>%  # Keep p_numeric for bolding condition
  gt() %>%
  tab_header(
    title = md("**Table X**"),
    subtitle = md("**Mann-Whitney U Test Results for Gesture Outcomes**")
  ) %>%
  cols_label(
    Outcome = "Outcome Variable",
    W = "W",
    p = "p",
    r = "Effect Size (r)"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows = p_numeric < 0.05
    )
  ) %>%
  tab_options(
    table.font.size = "small",
    table.align = "center",
    data_row.padding = px(2)
  ) %>%
  cols_hide(columns = p_numeric)  # Hide helper column p_numeric

```

## old Mann-Whtiney for rep gest
```{r, eval=FALSE}
# Mann-Whitney U Test (nonparametric equivalent to independent-samples t-test)
wilcox.test(rep_gestures ~ groupstatus, data = pp_gesture_data_ALLyears)

#report medians (which is preferred for nonparametric tests)
pp_gesture_data_ALLyears %>%
  group_by(groupstatus) %>%
  summarise(
    Median_RepGestures = median(rep_gestures, na.rm = TRUE),
    n = n()
  )

# Step 2a: Get N
n1 <- pp_gesture_data_ALLyears %>%
  filter(groupstatus == "TD") %>%
  nrow()

n2 <- pp_gesture_data_ALLyears %>%
  filter(groupstatus == "PL") %>%
  nrow()

N <- n1 + n2

# Step 2b: Approximate Z from p-value
# Two-tailed p-value, so divide by 2
z_approx <- qnorm(mw_result$p.value / 2, lower.tail = FALSE)

# Step 2c: Calculate r
r_effect_size <- z_approx / sqrt(N)

# Print it
r_effect_size



```

## boxplot of rep gest
```{r, eval=FALSE}
library(ggplot2)

ggplot(pp_gesture_data_ALLyears, aes(x = groupstatus, y = rep_gestures)) +
  
  # Boxplot
  geom_boxplot(fill = "gray80", color = "black", width = 0.6) +
  
  # Add individual data points
  geom_jitter(width = 0.1, height = 0, alpha = 0.5, size = 2) +
  
  # Labels
  labs(
    title = "Representational Gestures by Group Status",
    x = "Group Status",
    y = "Number of Representational Gestures"
  ) +
  
  # APA-style clean theme
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    legend.position = "none"
  )

```

## split bar graph
```{r, eval=FALSE}
# Create binary variable: 0 = No rep gesture, 1 = At least one rep gesture
pp_gesture_data_ALLyears <- pp_gesture_data_ALLyears %>%
  mutate(rep_gesture_present = if_else(rep_gestures > 0, 1, 0))

# Bar plot
ggplot(pp_gesture_data_ALLyears, aes(x = groupstatus, fill = factor(rep_gesture_present))) +
  geom_bar(position = "fill", color = "black") +
  scale_fill_manual(
    values = c("gray80", "black"),
    labels = c("Absent", "Present"),
    name = "Representational Gesture"
  ) +
  labs(
    title = "Proportion of Children Producing Representational Gestures",
    x = "Group Status",
    y = "Proportion"
  ) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.border = element_rect(color = "black", fill = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
#Figure 1. Proportion of children producing representational gestures during pretend play, by group status.```
```

## if want figure with individual measures
```{r, eval=FALSE}
ggplot(pp_gesture_data_ALLyears, aes(x = groupstatus, y = rep_gestures)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.7, size = 2, color = "black") +
  labs(
    title = "Representational Gesture Counts by Group",
    x = "Group Status",
    y = "Number of Representational Gestures"
  ) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.border = element_rect(color = "black", fill = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )
#Figure 2. Individual participant counts of representational gestures during pretend play, grouped by group status.
```

## treat rep gestures as binary
```{r, eval=FALSE}
# Make sure rep_gesture_present is coded 0/1 (binary)
pp_gesture_data_ALLyears <- pp_gesture_data_ALLyears %>%
  mutate(rep_gesture_present = if_else(rep_gestures > 0, 1, 0))

# Create a table
table_group_gesture <- table(pp_gesture_data_ALLyears$groupstatus, pp_gesture_data_ALLyears$rep_gesture_present)

# Run chi-square test
chisq.test(table_group_gesture)
```

## split bargraph
```{r, eval=FALSE}
library(ggplot2)
library(dplyr)

# Make sure rep_gesture_present is binary (0 = Absent, 1 = Present)
pp_gesture_data_ALLyears <- pp_gesture_data_ALLyears %>%
  mutate(rep_gesture_present = if_else(rep_gestures > 0, 1, 0))

# Create the bar plot
ggplot(pp_gesture_data_ALLyears, aes(x = groupstatus, fill = factor(rep_gesture_present))) +
  
  geom_bar(position = "fill", color = "black") +
  
  scale_fill_manual(
    values = c("gray80", "black"),  # APA-style colors
    labels = c("Absent", "Present"),  # Legend labels
    name = "Representational Gesture Presence"
  ) +
  
  labs(
    title = "Proportion of Children Producing Representational Gestures",
    x = "Group Status",
    y = "Proportion"
  ) +
  
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  
  theme(
    panel.border = element_rect(color = "black", fill = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    axis.text = element_text(color = "black")
  )

```

## 1. Conduct ANOVAs for gestures & summarize results
```{r, eval=FALSE}
# For total gestures
total_anova <- aov(total_gestures ~ groupstatus, pp_gesture_data_ALLyears)
cat("\nTotal gestures ANOVA:\n")
print(summary(total_anova))

# For representational gestures
rep_anova <- aov(rep_gestures ~ groupstatus, data = pp_gesture_data_ALLyears)
cat("\nRepresentational gestures ANOVA:\n")
print(summary(rep_anova))
 
# For proportion of representational gestures to total gestures
prop_rep_anova <- aov(prop_rep ~ groupstatus, data = pp_gesture_data_ALLyears)
cat("\nProportion of representational gestures ANOVA:\n")
print(summary(prop_rep_anova))
```

## 2. Calculate effect sizes
```{r, eval=FALSE}
if(require(effectsize)) {
  cat("\nEffect sizes:\n")
  cat("Total gestures: ")
  print(eta_squared(total_anova))
  cat("Proportion representational: ")
  print(eta_squared(prop_rep_anova))
} else {
  # Manual calculation if package not available
  cat("\nEffect sizes calculated manually:\n")
  # Formula for eta-squared: SS_between / SS_total
  summary_total <- summary(total_anova)
  eta_sq_total <- summary_total[[1]]["GroupStatus", "Sum Sq"] / 
                 sum(summary_total[[1]][, "Sum Sq"])
  cat("Total gestures eta-squared: ", eta_sq_total, "\n")
  # Repeat for other ANOVAs
}
```

## 3. Descriptive statistics by group
```{r, eval=FALSE}
group_stats <- pp_gesture_data_ALLyears %>%
  group_by(groupstatus) %>%
  summarize(
    n = n(),
    total_mean = mean(total_gestures, na.rm = TRUE),
    total_sd = sd(total_gestures, na.rm = TRUE),
    rep_mean = mean(rep_gestures, na.rm = TRUE),
    rep_sd = sd(rep_gestures, na.rm = TRUE),
    prop_rep_mean = mean(prop_rep, na.rm = TRUE),
    prop_rep_sd = sd(prop_rep, na.rm = TRUE)
  )

print("Descriptive statistics by group:")
print(group_stats)
```

## 4. Visualization of results
```{r, eval=FALSE}
if(require(tidyverse)) {
  # Raw counts visualization
  counts_long <- pp_gesture_data_ALLyears %>%
    select(groupstatus, total_gestures, rep_gestures) %>%
    pivot_longer(cols = c(total_gestures, rep_gestures),
                names_to = "gesture_type", 
                values_to = "count")
  
  p1 <- ggplot(counts_long, aes(x = gesture_type, y = count, fill = groupstatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.9)) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.9)) +
    labs(title = "Mean Gesture Counts During Pretend Play by Group",
         x = "Gesture Type",
         y = "Mean Count",
         fill = "Group Status") +
    theme_minimal() +
    scale_x_discrete(labels = c("total_gestures" = "Total",
                               "rep_gestures" = "Representational"))
  print(p1)
  
  # Proportions visualization
  props_long <- pp_gesture_data_ALLyears %>%
    select(groupstatus, prop_rep) %>%
    pivot_longer(cols = c(prop_rep),
                names_to = "proportion_type", 
                values_to = "proportion")
  
  p2 <- ggplot(props_long, aes(x = proportion_type, y = proportion, fill = groupstatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.9)) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.9)) +
    labs(title = "Proportion of Gesture Types During Pretend Play by Group",
         x = "Proportion Type",
         y = "Mean Proportion",
         fill = "Group Status") +
    theme_minimal() +
    scale_x_discrete(labels = c("prop_rep" = "Representational/Total")) +
    scale_y_continuous(labels = scales::percent)
  print(p2)
}
```

## 5. Statistical tests to directly compare TD and BI groups
```{r, eval=FALSE}
# t-tests for each variable (alternative to ANOVA with only two groups)
t_total <- t.test(total_gestures ~ groupstatus, data = pp_gesture_data_ALLyears)
t_rep <- t.test(rep_gestures ~ groupstatus, data = pp_gesture_data_ALLyears)
t_prop_rep <- t.test(prop_rep ~ groupstatus, data = pp_gesture_data_ALLyears)

cat("\nt-test results (direct comparison between groups):\n")
cat("\nTotal gestures:\n")
print(t_total)
cat("\nRepresentational gestures:\n")
print(t_rep)
cat("\nProportion representational:\n")
print(t_prop_rep)
```

## 6. Calculate Cohen's d effect sizes for t-tests with error handling
```{r, eval=FALSE}

cat("\nCohen's d effect sizes:\n")

# Function to safely calculate Cohen's d
safe_cohens_d <- function(formula, data) {
  tryCatch({
    res <- cohens_d(formula, data = data)
    return(res)
  }, error = function(e) {
    # Extract variable name from formula
    var_name <- as.character(formula)[2]
    # Get means by group
    means <- aggregate(formula, data = data, FUN = mean, na.rm = TRUE)
    sds <- aggregate(formula, data = data, FUN = sd, na.rm = TRUE)
    
    cat("Error calculating Cohen's d for", var_name, ":\n")
    cat("Group means:", toString(means), "\n")
    cat("Group SDs:", toString(sds), "\n")
    cat("Error message:", e$message, "\n")
    return(NULL)
  })
}

# Apply the safe function to each variable
cat("\nTotal gestures:\n")
print(safe_cohens_d(total_gestures ~ groupstatus, data = pp_gesture_data_ALLyears))

cat("\nRepresentational gestures:\n")
print(safe_cohens_d(rep_gestures ~ groupstatus, data = pp_gesture_data_ALLyears))

cat("\nProportion representational:\n")
print(safe_cohens_d(prop_rep ~ groupstatus, data = pp_gesture_data_ALLyears))
```

## Make APA-formatted data visualizations
```{r, eval=FALSE}
if(require(tidyverse)) {
  # Set APA theme
  apa_theme <- theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(colour = "black", fill = NA),
      axis.text = element_text(colour = "black", size = 12),
      axis.title = element_text(size = 12, face = "bold"),
      legend.title = element_text(size = 12, face = "bold"),
      legend.text = element_text(size = 12),
      legend.position = "bottom",
      legend.background = element_rect(fill = "white", colour = "black"),
      plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
      text = element_text(family = "Times New Roman")
    )
  
  # Raw counts visualization
  counts_long <- pp_gesture_data_ALLyears %>%
    select(groupstatus, total_gestures, rep_gestures) %>%
    pivot_longer(cols = c(total_gestures, rep_gestures),
                names_to = "gesture_type", 
                values_to = "count")
  
  # Renaming GroupStatus labels to be more readable
  counts_long$groupstatus <- factor(counts_long$groupstatus, 
                                  levels = c("PL", "TD"),
                                  labels = c("Perinatal Lesions", "Typically Developing"))
  
  p1 <- ggplot(counts_long, aes(x = gesture_type, y = count, fill = groupstatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.8), 
                alpha = 0.8, color = "black", size = 0.2) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.8)) +
    labs(title = "Figure 1",
         subtitle = "Mean Gesture Counts During Pretend Play by Group",
         x = "Gesture Type",
         y = "Mean Count",
         fill = "Group") +
    scale_x_discrete(labels = c("total_gestures" = "Total",
                               "rep_gestures" = "Representational")) +
    scale_fill_grey(start = 0.4, end = 0.8) +
    apa_theme +
    theme(plot.subtitle = element_text(size = 12, hjust = 0.5))
  
  print(p1)
  
  # Save high-resolution figure
  ggsave("Figure_1_Gesture_Counts.tiff", p1, width = 7, height = 5, dpi = 300)
  
  # Proportions visualization
  props_long <- pp_gesture_data_ALLyears %>%
    select(groupstatus, prop_rep) %>%
    pivot_longer(cols = c(prop_rep),
                names_to = "proportion_type", 
                values_to = "proportion")
  
  # Renaming GroupStatus labels
  props_long$groupstatus <- factor(props_long$groupstatus, 
                                  levels = c("PL", "TD"),
                                  labels = c("Perinatal Lesions", "Typically Developing"))
  
  p2 <- ggplot(props_long, aes(x = proportion_type, y = proportion, fill = groupstatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.8), 
                alpha = 0.8, color = "black", size = 0.2) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.8)) +
    labs(title = "Figure 2", 
         subtitle = "Proportion of Gesture Types During Pretend Play by Group",
         x = "Proportion Type",
         y = "Mean Proportion",
         fill = "Group") +
    scale_x_discrete(labels = c("prop_rep" = "Representational/Total")) +
    scale_y_continuous(labels = scales::percent, 
                      breaks = seq(0, 0.25, 0.05)) +
    scale_fill_grey(start = 0.4, end = 0.8) +
    apa_theme +
    theme(plot.subtitle = element_text(size = 12, hjust = 0.5))
  
  print(p2)
  
  # Save high-resolution figure
  ggsave("Figure_2_Gesture_Proportions.tiff", p2, width = 7, height = 5, dpi = 300)
}
```

## Is rep_gestures positively correlated with total_gestures (overall gestures produced during pretend play)?
```{r, eval=FALSE}
cor.test(pp_gesture_data_ALLyears$rep_gestures, pp_gesture_data_ALLyears$total_gestures)

```

## boxplot of representational gesture presence & narrative score
```{r, eval=FALSE}
# First: recode rep_gestures as a factor for clean labeling
pp_gesture_data_ALLyears <- pp_gesture_data_ALLyears %>%
  mutate(rep_gesture_factor = factor(rep_gesture_present, levels = c(0,1), labels = c("Absent", "Present")))

# Plot
ggplot(pp_gesture_data_ALLyears, aes(x = rep_gesture_factor, y = max_avg)) +
  geom_boxplot(fill = "gray80", color = "black") +
  stat_summary(fun = mean, geom = "point", size = 3, color = "black", position = position_dodge(width=0.75)) +
  labs(
    x = "Representational Gesture",
    y = "Narrative Structure Score",
    title = "Narrative Structure by Representational Gesture Presence"
  ) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.border = element_rect(color = "black", fill = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold")
  )

```

```{r, eval=FALSE}
library(ggplot2)
library(dplyr)

# Recode rep_gestures as a factor for clean group labels
pp_gesture_data_ALLyears <- pp_gesture_data_ALLyears %>%
  mutate(rep_gesture_factor = factor(rep_gesture_present, levels = c(0, 1),
                                     labels = c("Absent", "Present")))

# Bar graph with error bars (mean ± SE)
ggplot(pp_gesture_data_ALLyears, aes(x = rep_gesture_factor, y = max_avg)) +
  
  # Bar for mean narrative scores
  stat_summary(fun = mean, geom = "bar", fill = "gray80", color = "black", width = 0.6) +
  
  # Error bars (± standard error)
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black", size = 0.8) +
  
  # Points for individual data points (optional, adds transparency)
  geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.5) +
  
  # Labels
  labs(
    title = "Narrative Structure by Representational Gesture Presence",
    x = "Representational Gesture Presence",
    y = "Narrative Structure Score"
  ) +
  
  # APA-style theme
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),
    axis.line = element_line(color = "black"),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )

#Figure 1
#Mean narrative structure scores for children who did versus did not produce representational gestures during pretend play. Error bars represent ±1 standard error of the mean. Children who produced representational gestures had marginally higher narrative structure scores compared to those who did not (p = .075).

```

```{r, eval=FALSE}
pp_gesture_data_ALLyears %>%
  group_by(rep_gesture_factor) %>%
  summarise(
    Mean_Narrative = mean(max_avg, na.rm = TRUE),
    SD_Narrative = sd(max_avg, na.rm = TRUE),
    n = n()
  )

```

```{r, eval=FALSE}
# Identify participants with missing rep_gestures
pp_gesture_data_ALLyears %>%
  filter(is.na(rep_gestures)) %>%
  select(participant_id, total_gestures, rep_gestures, everything())

# List all unique values
unique(pp_gesture_data_ALLyears$rep_gestures)

# Find participants with rep_gestures not equal to 0 or 1
pp_gesture_data_ALLyears %>%
  filter(!(rep_gestures %in% c(0, 1))) %>%
  select(participant_id, rep_gestures)

```

