---
title: "Summary Stats"
author: "Kristen Johnson"
date: "2025-02-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary libraries
library(readxl)   # For reading Excel files
library(dplyr)    # For binding and data manipulation
library(stringr) # For string functions
library(lubridate)  # For handling datetime columns (if needed)
```


```{r, echo=FALSE}
# Read in data
csv_all <- read.csv("/Users/kristenjohnson/KristenWorkingDirectory/Narrative Data/CSVs_of_Combined_Data/P3_H8_ALL.csv")
```
## Summary Statistics for P2_H8

### ALL Gestures by Participant ID
```{r, echo=FALSE}

# Count the number of times individual participants made any gesture
gesture_count <- csv_all %>%
  group_by(participant_id) %>% # Groups data by participant ID
  summarise(gesture_all = sum(!is.na(c_form))) # Counts number of non NA values

print(gesture_count)
```

### Total number of gestures among all observations of all participants:
```{r, echo=FALSE}
# Sum of all instances of gesture as counted in the non NA values in the above data frame
total_gestures <- sum(gesture_count$gesture_all)

print(total_gestures)

```



### Mean and Standard deviation of ALL gestures 
```{r, echo=FALSE}
# get summary statistics of the count_non_na column of the above dataset
gesture_summary <- gesture_count %>% 
  summarise(
    mean_gesture = mean(gesture_all),
    sd_gesture = sd(gesture_all)
  )

print(gesture_summary)

```


### A Data Frame of only iconic gestures. Summary Stats to follow.
```{r, echo=FALSE}
# Load necessary libraries
#library(dplyr)   # For data manipulation
#library(stringr) # For string functions

# Create a new data frame that includes rows where 'iconic' appears anywhere in c_form
iconic_df <- csv_all %>%
  filter(str_detect(str_to_lower(c_form), "iconic"))

# View the first few rows of the new data frame
head(iconic_df)
```

### Number of iconic gestures per participant ID 
```{r, echo=FALSE}
# Count occurrences of 'iconic' for each participant_id
iconic_counts <- iconic_df %>%
  group_by(participant_id) %>%
  summarise(iconic_count = sum(str_detect(str_to_lower(c_form), "iconic"), na.rm = TRUE), .groups = "drop")

# Print results
print(iconic_counts)
```
### Total number of iconic gestures across all participants:
```{r, echo=FALSE}
# The sum of all instances of iconic gestures in the above column
all_iconic <- sum(iconic_counts$iconic_count)

print(all_iconic)
```



### Summary statistics of iconic gestures across participants
```{r, echo=FALSE}
# Calculate mean and standard deviation of 'iconic' counts across participants
iconic_summary <- iconic_counts %>%
  summarise(
    mean_iconic = mean(iconic_count, na.rm = TRUE),
    sd_iconic = sd(iconic_count, na.rm = TRUE)
  )

# Print results
print(iconic_summary)
```
### Instances of Pretend Play by Participant ID:
```{r, echo=FALSE}
# Count the number of times individual participants engaged in pretend play
pp_count <- csv_all %>%
  group_by(participant_id) %>%  # Groups data by participant ID
  summarise(
    count_p_combined = sum(c_pret, na.rm = TRUE) # for the c_pret column name
    #count_p_combined = sum(C_Pret, na.rm = TRUE) # for the C_Pret column name
    #count_p_combined = sum(str_detect(str_to_lower(`Child DXT`), "p"), na.rm = TRUE) 
    # In at least one dataset there are two Child DXT columns thanks to one file not capitalizing it. if this is the case, comment out the above line and remove the comment hash in front of the code below
    #count_p_combined = sum(str_detect(str_to_lower(`Child DXT`), "p"), na.rm = TRUE) + sum(str_detect(str_to_lower(`CHILD DXT`), "p"), na.rm = TRUE)  
  )

print(pp_count)
```

### Total number of instances of pretend play:
```{r, echo=FALSE}
# Sum the coun_p_combined column from above
all_pp <- sum(pp_count$count_p_combined)

print(all_pp)
```

### Descriptive statistics of all instances of pretend play:
```{r, echo=FALSE}
# calculate the summary statistics of count_p_combined
pp_summary <- pp_count %>%
  summarise(
    mean_pp = mean(count_p_combined, na.rm = TRUE),
    sd_pp = sd(count_p_combined, na.rm = TRUE)
  )

# Print results
print(pp_summary)
```

### Average words per utterance by Participant ID
```{r, echo=FALSE}
# finding the average words per utterance per participant
avg_words_per_utt <- csv_all %>%
  group_by(participant_id) %>%
  summarise(
    total_words = sum(C_wpu[C_wpu > 0], na.rm = TRUE),  # Sum of non-zero word counts
    total_utts = sum(C_wpu > 0, na.rm = TRUE),  # Count of non-zero utterances
    avg_c_wpu = ifelse(total_utts > 0, total_words / total_utts, NA)  # Calculate mean, avoid divide by zero
  )

print(avg_words_per_utt)

```

### Adding in Standar Deviation to the above table
```{r, echo=FALSE}
# finding the standard deviation for words per utterance per participant
avg_words_per_utt <- csv_all %>%
  group_by(participant_id) %>%
  summarise(
    total_words = sum(C_wpu[C_wpu > 0], na.rm = TRUE),  # Sum of non-zero word counts
    total_utts = sum(C_wpu > 0, na.rm = TRUE),  # Count of non-zero utterances
    avg_c_wpu = ifelse(total_utts > 0, total_words / total_utts, NA),  # Mean words per utterance
    sd_c_wpu = ifelse(total_utts > 1, sd(C_wpu[C_wpu > 0], na.rm = TRUE), NA)  # Standard deviation
  )

print(avg_words_per_utt)

```

### Mean of Means in Above table
```{r, echo=FALSE}
# calculate the summary statistics of count_p_combined
c_wpu_summary <- avg_words_per_utt %>%
  summarise(
    c_total_words = sum(total_words, na.rm = TRUE),
    mean_c_wpu = mean(avg_c_wpu, na.rm = TRUE),
    sd_c_wpu = sd(avg_c_wpu, na.rm = TRUE)
  )

# Print results
print(c_wpu_summary)
```
### Mean of Avg Words per Utternace by total number of words over total number of utternaces
```{r, echo=FALSE}

tw_o_tu <- sum(avg_words_per_utt$total_words)/sum(avg_words_per_utt$total_utts)
print(tw_o_tu)
```

### count of instances when gesture and pretend play occur simulatneously

