---
title: "PN_Demographic-Data"
author: "Kristen Johnson"
date: "2025-03-08"
output: html_document
---

SETUP
```{r, echo=FALSE}
# Install the gt package if not already installed
#install.packages("gt")
# Install and load the flextable package
#install.packages("flextable")
# Install and load kableExtra
#install.packages("kableExtra")

# load necessary packages
library(knitr)  ##Load the knitr package for kable
library(gt) ## Load the gt package
library(flextable) ##load the flextable package
library(kableExtra) ##load the kableExtra package
library(here)
```

# P3 Children's Neurological Data for P3 children 
-recorded ages from when they participated in Stein's Story Stem Task narrative measure during Literacy Visits
```{r, echo=FALSE}
# Create a dataframe for Neurological Data of P3
P3_neurological_data <- data.frame(
  ID = c(30, 35, 46, 93, 94, 98, 99, 117, 132, 135, 150),
  Age_at_visit = c(5.52, 6.33, 5.82, 5.65, 6.32, 8.96, 7, 5.96, 5.64, 5.13, 5.25),
  Sex = c("Female", "Female", "Female", "Male", "Female", "Male", "Male", "Female", "Female", "Female", "Female"),
  Side = c("Left", "Left", "Right", "Right", "Right", "Left", "Left", "Right", "Left", "Right", "Left"),
  Size = c("Large", "Medium", "Large", "Small", "Small", "Small", "Large", "Medium", "Small", "Small", "Small"),
  Type = c("Cerebrovascular Infarct", "Periventricular", "Cerebrovascular Infarct", "Periventricular", "Periventricular", "Cerebrovascular Infarct", "Cerebrovascular Infarct", "Cerebrovascular Infarct", "Periventricular", "Cerebrovascular Infarct", "Periventricular"),
  Seizure = c("No", "No", "Yes", "Yes", "No", "Yes", "Yes", "Yes", "No", "No*", "No"),
  Areas_affected = c(
    "Frontal, Temporal, Parietal, Occipital, Subcortical", "Subcortical", "Frontal, Temporal, Parietal, Subcortical", "Subcortical",
    "Temporal, Parietal, Internal Capsule, Subcortical", "Frontal, Temporal, Subcortical", "Temporal, Parietal, Occipital, Internal Capsule, Subcortical", "Frontal, Temporal, Parietal, Subcortical", "Temporal, Subcortical", "Frontal, Parietal", "White Matter, Subcortical"
  ),
  Premature = c("No", "36.5 wk", "34 wk", "No", "No", "No", "No", "No", "No", "No", "No")
)

# Print the dataframe
print(P3_neurological_data)
```
Note: Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication.

```{r, echo=FALSE}
# Save the dataframe to a CSV file
write.csv(P3_neurological_data, here("PN_Datasets", "Demographics", "P3_neurological_data.csv", row.names = FALSE))
```

```{r}
neurological_data <- read.csv(here("PN_Datasets", "Demographics", "P3_neurological_data.csv"))
```

#Attempts at tables
## Simple Table Using knittr::kable()
```{r, echo=FALSE}

# Create a table with a note to then print the table in a well-formatted way
kable(neurological_data, caption = "Neurological Information about Children with Pre- or Perinatal Brain Injury") %>%
  footnote(general = "Codes are Type (CI, cerebrovascular infarct; Pv, periventricular), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication."
          ) 
```

## More Customizable Table using gt::gt()
```{r,echo=FALSE}
# Create and format the table
neurological_data %>%
  gt() %>%
  tab_header(
    title = "Neurological Information about Children",
    subtitle = "Pre- or Perinatal Brain Injury"
  ) %>%
  tab_footnote(
    footnote = "Note: F=frontal, T=temporal, P=parietal, O,=occipital IC=internal capsule (white matter in the frontal area), WM=white matter; * Neonatal seizures resolved without medication."
  ) %>%
  fmt_number(
    columns = "Age_at_visit",
    decimals = 2
  ) %>%
  cols_label(
    ID = "ID",
    Age_at_visit = "Age (yrs)",
    Sex = "Sex",
    Side = "Side",
    Size = "Size",
    Type = "Type",
    Seizure = "Seizure History",
    Areas_affected = "Affected Areas",
    Premature = "Premature Status"
  )
```

## Word-Doc-Ready Table using flextable
```{r, echo=FALSE}
# Create a nicely formatted table
flextable(neurological_data) %>%
  set_caption("Neurological Information about Children with Pre- or Perinatal Brain Injury") %>%
  add_footer_lines("Note: Codes are Type (CI, cerebrovascular infarct; Pv, periventricular), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter): * Neonatal seizures resolved without medication.") %>%
  autofit()
```

## APA-style table for Quarto document
```{r, echo=FALSE}
# Create an APA-style table with a caption
kable(neurological_data, booktabs = TRUE, digits = 2,
      caption = "Neurological Information about Children with Pre- or Perinatal Brain Injury") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  footnote(general = "Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication."
           ) 
```

_____________________________________________________________________________

# Demographic Data for Groups P2 & P3
```{r}
Play_Narrative_Demographics_xl <- read_xlsx(here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Play_Narrative_Demographics.xlsx")) %>%
  mutate_all(~ifelse(is.numeric(.), round(., 2), .))

# Replace "NoData" with NA
Play_Narrative_Demographics_xl$FatherEdYrs07[Play_Narrative_Demographics_xl$FatherEdYrs07 == "NoData"] <- NA

# Convert to numeric
Play_Narrative_Demographics_xl$FatherEdYrs07 <- as.numeric(Play_Narrative_Demographics_xl$FatherEdYrs07)

# Convert to numeric
Play_Narrative_Demographics_xl$MotherEdYrs07 <- as.numeric(Play_Narrative_Demographics_xl$MotherEdYrs07)

# save it to a new file for future use:
write.csv(Play_Narrative_Demographics, here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N","Play_Narrative_Demographics.csv", row.names = FALSE))
```

# Extract specific data for in line references in Qmd
```{r}
# read in demogrpahics for both P3 and P2
Play_Narrative_Demographics <- read_xlsx(here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Play_Narrative_Demographics.xlsx"))

# Function to calculate percentages
calculate_percentage <- function(data, variable) {
  data %>%
    count(!!sym(variable)) %>%
    mutate(Percentage = round(n / sum(n) * 100, 1))
}

# Filter data for Group 2 and Group 3
demographics_group2 <- Play_Narrative_Demographics %>% filter(Group == 2)
demographics_group3 <- Play_Narrative_Demographics %>% filter(Group == 3)

# Compute percentages
gender_group2 <- calculate_percentage(demographics_group2, "ChildGender")
gender_group3 <- calculate_percentage(demographics_group3, "ChildGender")

race_group2 <- calculate_percentage(demographics_group2, "ChildRace")
race_group3 <- calculate_percentage(demographics_group3, "ChildRace")

ethnicity_group2 <- calculate_percentage(demographics_group2, "ChildEthnicity")
ethnicity_group3 <- calculate_percentage(demographics_group3, "ChildEthnicity")
```

