---
title: "Demographic-Data_P3"
author: "Kristen Johnson"
date: "2025-03-08"
output: html_document
---

```{r, echo=FALSE}
# Install the gt package if not already installed
install.packages("gt")
# Install and load the flextable package
install.packages("flextable")
# Install and load kableExtra
install.packages("kableExtra")


# load necessary packages
library(knitr)  ##Load the knitr package for kable
library(gt) ## Load the gt package
library(flextable) ##load the flextable package
library(kableExtra) ##load the kableExtra package
```


```{r, echo=FALSE}
P3_demographics <- data.frame()
```

# P3 Children's Neurological Data for P3 children 
-recorded ages from when they participated in Stein's Story Stem Task narrative measure during Literacy Visits
```{r, echo=FALSE}
# Create a dataframe for Neurological Data of P3
P3_neurological_data <- data.frame(
  ID = c(30, 35, 46, 93, 94, 98, 99, 117, 132, 135, 150),
  Age_at_visit = c(5.52, 6.33, 5.82, 5.65, 6.32, 8.96, 7, 5.96, 5.64, 5.13, 5.25),
  Sex = c("F", "F", "F", "M", "F", "M", "M", "F", "F", "F", "F"),
  Side = c("L", "L", "R", "R", "R", "L", "L", "R", "L", "R", "L"),
  Size = c("L", "M", "L", "S", "S", "S", "L", "M", "S", "S", "S"),
  Type = c("CI", "Pv", "CI", "Pv", "Pv", "CI", "CI", "CI", "Pv", "CI", "Pv"),
  Seizure = c("No", "No", "Yes", "Yes", "No", "Yes", "Yes", "Yes", "No", "No*", "No"),
  Areas_affected = c(
    "F, T, P, O, subcortical", "Subcortical", "F, T, P, subcortical", "Subcortical",
    "T, P, IC, subcortical", "F, T, subcortical", "T, P, O, IC, subcortical",
    "F, T, P, subcortical", "T, subcortical", "F, P", "WM, subcortical"
  ),
  Premature = c("No", "36.5 wk", "34 wk", "No", "No", "No", "No", "No", "No", "No", "No")
)

# Print the dataframe
print(P3_neurological_data)
```
Note: Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication.

```{r, echo=FALSE}
# Save the dataframe to a CSV file
write.csv(P3_neurological_data, "P3_neurological_data.csv", row.names = FALSE)
```

## Simple Table Using knittr::kable()
```{r, echo=FALSE}

# Create a table with a note to then print the table in a well-formatted way
kable(neurological_data, caption = "Neurological Information about Children with Pre- or Perinatal Brain Injury") %>%
  footnote(general = "Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication."
          ) 
```

## More Customizable Table using gt::gt()
```{r,echo=FALSE}
# Create and format the table
neurological_data %>%
  gt() %>%
  tab_header(
    title = "Neurological Information about Children",
    subtitle = "Pre- or Perinatal Brain Injury"
  ) %>%
  tab_footnote(
    footnote = "Note: Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication."
  ) %>%
  fmt_number(
    columns = "Age_at_visit",
    decimals = 2
  ) %>%
  cols_label(
    ID = "ID",
    Age_at_visit = "Age (yrs)",
    Sex = "Sex",
    Side = "Side",
    Size = "Size",
    Type = "Type",
    Seizure = "Seizure History",
    Areas_affected = "Affected Areas",
    Premature = "Premature Status"
  )
```

## Word-Doc-Ready Table using flextable
```{r, echo=FALSE}
# Create a nicely formatted table
flextable(neurological_data) %>%
  set_caption("Neurological Information about Children with Pre- or Perinatal Brain Injury") %>%
  add_footer_lines("Note: Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication.") %>%
  autofit()
```

APA-style table for Quarto document
```{r, echo=FALSE}
# Create an APA-style table with a caption
kable(neurological_data, booktabs = TRUE, digits = 2,
      caption = "Neurological Information about Children with Pre- or Perinatal Brain Injury") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  footnote(general = "Codes are Sex (F, female; M, male), Side (L, left; R, right), Size (S, small; M, medium; L, large), Type (CI, cerebrovascular infarct; Pv, periventricular), Seizure (Y. history of seizures; N, no history), Areas affected (F, frontal; T, temporal; P, parietal; O, occipital; IC, internal capsule (white matter in the frontal area); WM, white matter); periventricular lesions involve the thalamus, basal ganglia, the medial temporal lobe and/or white matter tracts. * Neonatal seizures resolved without medication."
           ) 
```

