---
title: "Narrative _Data_Wrangling"
author: "Kristen Johnson"
date: "2025-03-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(stringr)
library(tidyverse)
library(here)
```
# Narrative scores and codes:
1. No structure (**NS**)- only one clause with or without a verb. 
2. Descriptive story (**DS**)- a series of clauses that do not report actions constrained by temporality.
3. Action story (**AC**)-  includes actions *with a temporal order*, but story events are not causally organized.
4. Reactive story (**RS**)- includes actions that are causally organized.
5. Incomplete goal-based story (**IGB**)- contains a goal statement, but no attempt or outcome following the goal.
6. Incomplete goal-based with attempt (**IGBA**)- contains a goal statement, and attempt(s), but not outcome following the goal.
7. Complete goal-based story with one episode implicit goal (**GB1-IG**)- includes temporal and causal structure, as well as *implicitly*-stated goal of the protagonist, *one* series of attempts to achieve the goal, and an outcome of these attempts.
8. Complete goal-based story with one episode explicit goal (**GB1-EG**)- includes temporal and causal structure, as well as *explicitly*-stated goal of the protagonist, *one* series of attempts to achieve the goal, and an outcome of these attempts.
9. Complete goal-based story with multiple episodes implicit goal (**GBM-IG**)- includes *multiple* goal-attempt-outcome sequences and frequently more than one *implicitly*-stated goal. 
10. Complete goal-based story with multiple episodes explicit goal (**GBM-EG**)- includes *multiple* goal-attempt-outcome sequences. with at least one of the goals stated *explicitly*.

# Load in data & convert narrative scores to numbers:
```{r}
# Load necessary libraries
library(dplyr)
library(readxl)

# Read in the data
Narrative_Results_Simple <- read_xlsx(here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Simple.xlsx"))

# Get column names of the data frame
column_names <- colnames(Narrative_Results_Simple)

# Print the column names
print(column_names)


# Define the abbreviation-to-number mapping
replacements <- c("NS" = 1, "DS" = 2, "AC" = 3, "RS" = 4, "IGB" = 5, "IGBA" = 6, "GB1-IG" = 7, "GB1-EG" = 8, "GBM-IG" = 9, "GBM-EG" = 10)

# Replace abbreviations in multiple columns and handle NA
Narrative_Results_Numeric <- Narrative_Results_Simple %>%
  mutate(across(c("FoxHOS07", "AliceHOS07", "AlanHOS07",  "FoxHOS08",  "AlanHOS08", "FoxHOS09", "AlanHOS09" ), ~ case_when(
    . == "NS" ~ 1,
    . == "DS" ~ 2,
    . == "AC" ~ 3,
    . == "RS" ~ 4,
    . == "IGB" ~ 5,
    . == "IGBA" ~ 6,
    . == "GB1-IG" ~ 7,
    . == "GB1-EG" ~ 8,
    . == "GBM-IG" ~ 9,
    . == "GBM-EG" ~ 10,
    TRUE ~ as.numeric(NA)  # In case of any unmatched value, set to NA
  ))) %>%
  mutate(across(c("FoxHOS07", "AliceHOS07", "AlanHOS07",  "FoxHOS08",  "AlanHOS08", "FoxHOS09", "AlanHOS09"), ~ replace_na(., 0)))  # Replace NA with 0 in the selected columns

# Write the updated data frame into a csv
write_csv(Narrative_Results_Numeric, here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Numeric.csv"))

```

# Determine each participant's highest Narrative Score in each year and overall Avg of high scores:
```{r}
# Collect highest score from each year
Narrative_Results_Score_Totals <- Narrative_Results_Numeric %>%
  rowwise() %>%
  mutate(
    HighScore07 = max(c_across(ends_with("07")), na.rm = TRUE),
    HighScore08 = max(c_across(ends_with("08")), na.rm = TRUE),
    HighScore09 = max(c_across(ends_with("09")), na.rm = TRUE),
    AvgHighScore = mean(c(HighScore07, HighScore08, HighScore09), na.rm = TRUE)  # Add average of the 3 high scores
  ) %>%
  ungroup()  # Remove rowwise grouping after mutation

write_csv(Narrative_Results_Score_Totals,  here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Score_Totals.csv"))
```

# Make new dataframe with relevant info only:
```{r}
# Make new datafram with relevant info only
Narrative_Results_Final <- Narrative_Results_Score_Totals %>%
  select(ChildID, Project, HighScore07, HighScore08, HighScore09, AvgHighScore)

print(Narrative_Results_Final)

write_csv(Narrative_Results_Final,  here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Final.csv"))

```

# Create new column with clearer group status for kids
```{r}
Narrative_Results_Final <- Narrative_Results_Final %>%
  mutate(GroupStatus = case_when(
    Project == 3 ~ "PL",
    Project == 2 ~ "TD",
    TRUE ~ NA_character_  # Ensures missing values are handled
  ))
```

# Renaming & grouping scores to simplify for categorical analysis
```{r}
# Categorize ability into Low, Medium, High
Cat_Narrative_Results_Final <- Narrative_Results_Final %>%
  mutate(ability_category = factor(
    case_when(
      AvgHighScore <= 3 ~ "Low",
      AvgHighScore <= 7 ~ "Medium",
      TRUE ~ "High"
    ),
    levels = c("Low", "Medium", "High")  # Ensuring consistent order
  ))

print(Cat_Narrative_Results_Final)
```

# Ensure a Specific Order for Grouping 
```{r}
#ensure BI kids appear first in plots and tables
Cat_Narrative_Results_Final <- Cat_Narrative_Results_Final %>%
  mutate(GroupStatus = fct_relevel(GroupStatus, "PL"))

print(Cat_Narrative_Results_Final)
```

# Check to see if ability categories align with actual scores
```{r}
# Order Ability Groups by Mean Narrative Score

Cat_Narrative_Results_Final <- Cat_Narrative_Results_Final %>%
  mutate(ability_category = fct_reorder(ability_category, AvgHighScore, .fun = mean))

```

# Boxplot of Narrative Scores by Group Status
```{r}
ggplot(Cat_Narrative_Results_Final, aes(x = GroupStatus, y = AvgHighScore, fill = GroupStatus)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Comparison of Narrative Ability Scores by Group Status",
       x = "Group Status",
       y = "Narrative Score")

```

# Bar Plot of Ability Categories by Group Status
```{r}
ggplot(Cat_Narrative_Results_Final, aes(x = ability_category, fill = GroupStatus)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Distribution of Narrative Ability Categories by Group Status",
       x = "Ability Level",
       y = "Count")

```

# Save it!
```{r}
write_csv(Cat_Narrative_Results_Final, "Narrative_Comparison.csv")
```

