---
title: "Narrative _Data_Wrangling"
author: "Kristen Johnson"
date: "2025-03-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(stringr)
```

1. No structure (**NS**)- only one clause with or without a verb. 
2. Descriptive story (**DS**)- a series of clauses that do not report actions constrained by temporality.
3. Action story (**AC**)-  includes actions *with a temporal order*, but story events are not causally organized.
4. Reactive story (**RS**)- includes actions that are causally organized.
5. Incomplete goal-based story (**IGB**)- contains a goal statement, but no attempt or outcome following the goal.
6. Incomplete goal-based with attempt (**IGBA**)- contains a goal statement, and attempt(s), but not outcome following the goal.
7. Complete goal-based story with one episode implicit goal (**GB1-IG**)- includes temporal and causal structure, as well as *implicitly*-stated goal of the protagonist, *one* series of attempts to achieve the goal, and an outcome of these attempts.
8. Complete goal-based story with one episode explicit goal (**GB1-EG**)- includes temporal and causal structure, as well as *explicitly*-stated goal of the protagonist, *one* series of attempts to achieve the goal, and an outcome of these attempts.
9. Complete goal-based story with multiple episodes implicit goal (**GBM-IG**)- includes *multiple* goal-attempt-outcome sequences and frequently more than one *implicitly*-stated goal. 
10. Complete goal-based story with multiple episodes explicit goal (**GBM-EG**)- includes *multiple* goal-attempt-outcome sequences. with at least one of the goals stated *explicitly*.

```{r}
#read in data for narrative outcomes
Narrative_Results_Simple <- read_xlsx("~/KristenWorkingDirectory/Play_Narrative/PN_Datasets/Narrative/CSVs_of_Combined_Data_N/Narrative_Results_Simple.xlsx")

# Define the abbreviation-to-number mapping
replacements <- c("NS" = 1, "DS" = 2, "AC" = 3, "RS" = 4, "IGB" = 5, "IGBA" = 6, "GB1-IG" = 7, "GB1-EG" = 8, "GBM-IG" = 9, "GBM-EG" = 10)

# Replace abbreviations, handle NA, and convert to numeric
Narrative_Results_Numeric <- Narrative_Results_Simple %>%
  mutate(across(everything(), ~ recode(., !!!replacements))) %>%  # Replace abbreviations
  mutate(across(everything(), ~ replace_na(., 0)))   # Replace NA with "0"
  
# write the updated data frame into a csv
write_csv(Narrative_Results_Numeric, "~/KristenWorkingDirectory/Play_Narrative/PN_Datasets/Narrative/CSVs_of_Combined_Data_N/Narrative_Results_Numeric.csv")
```

```{r}
# Determine each participant's overall Narrative Score

# Collect highest score from each year
Narrative_Results_Score_Totals <- Narrative_Results_Numeric %>%
  rowwise() %>%
  mutate(
    HighScore07 = max(c_across(ends_with("07")), na.rm = TRUE),
    HighScore08 = max(c_across(ends_with("08")), na.rm = TRUE),
    HighScore09 = max(c_across(ends_with("09")), na.rm = TRUE),
    AvgHighScore = mean(c(HighScore07, HighScore08, HighScore09), na.rm = TRUE)  # Add average of the 3 high scores
  ) %>%
  ungroup()  # Remove rowwise grouping after mutation

write_csv(Narrative_Results_Score_Totals, "~/KristenWorkingDirectory/Play_Narrative/PN_Datasets/Narrative/CSVs_of_Combined_Data_N/Narrative_Results_Score_Totals.csv")
```


```{r}
# Make new datafram with relevant info only
Narrative_Results_Final <- Narrative_Results_Score_Totals %>%
  select(ChildID, Project, HighScore07, HighScore08, HighScore09, AvgHighScore)

print(Narrative_Results_Final)

write_csv(Narrative_Results_Final, "~/KristenWorkingDirectory/Play_Narrative/PN_Datasets/Narrative/CSVs_of_Combined_Data_N/Narrative_Results_Final.csv")

```

