---
title: "PP_Narrative_Analysis_H10"
author: "Kristen Johnson"
date: "2025-04-04"
output:
  pdf_document: default
  html_document: default
---

```{r}
# install necessary packages

#install.packages("lmerTest")

# Load required packages
library(tidyverse)
library(car)       # For regression diagnostics
library(lmtest)    # For regression diagnostics
library(effectsize) # For effect sizes
library(ggplot2)   # For visualization
library(here)
library(dplyr)
library(lme4)
library(lmerTest)
library(extrafont)
```

```{r}
# find  project root directory automatically
merged_H10 <- read.csv(here("PN_Datasets", "Pretend_Play", "CSVs_of_Combined_Data_PP", "merged_H10.csv"))

H10_gesture_play <- merged_H10 %>%
  group_by(participant_id) %>%
  summarise(
    total_pretend_play = sum(c_pret, na.rm = TRUE),
    mlu = mean(C_wpu, na.rm = TRUE),
    any_pretend_play = if_else(any(c_pret == 1, na.rm = TRUE), 1, 0),
    total_rep_gestures = sum(gesture_rep == 1, na.rm = TRUE),
    total_all_gestures = sum(gesture_all == 1, na.rm = TRUE),
    group_status = first(GroupStatus)
  )

write_csv(H10_gesture_play, 
          here("PN_DataAnalysis", "PP_Narrative_Analysis", "H10_gesture_play.csv"))
```


```{r}

# read in data frame that is just gesture during instance of pretend play
pretend_data_H10 <- read.csv(here("PN_Datasets", "Pretend_Play", "CSVs_of_Combined_Data_PP","pretend_data_H8.csv"))

# Aggregate by child
child_summary_H10 <- pretend_data_H10 %>%
  group_by(participant_id, GroupStatus) %>% 
  summarize(
    total_gestures = sum(gesture_all),
    rep_gestures = sum(gesture_rep),
    icon_gestures = sum(gesture_icon),
    prop_rep = sum(gesture_rep) / sum(gesture_all),
    prop_icon = sum(gesture_icon) / sum(gesture_all),
  )
```

```{r}
# add mlu from all utterances 

pretend_data_H10 <- child_summary_H10 %>%
  left_join(select(H10_gesture_play, participant_id, mlu), by = "participant_id") %>%
  left_join(select(H10_gesture_play, participant_id, total_pretend_play), by = "participant_id")

write_csv(pretend_data_H10, here("PN_DataAnalysis", "PP_Narrative_Analysis", "pretend_data_H10.csv"))

```

```{r}
# Create binary variable for representational gesture presence
pretend_data_H10 <- pretend_data_H10 %>%
  mutate(rep_gesture_present = ifelse(rep_gestures > 0, 1, 0))

# Check distributions
hist(pretend_data_H10$total_gestures)
hist(pretend_data_H10$rep_gestures)
table(pretend_data_H10$rep_gesture_present, pretend_data_H10$GroupStatus)

# Check correlation between total gestures and representational gestures
cor.test(pretend_data_H10$total_gestures, pretend_data_H10$rep_gestures)
```

```{r}
# 2. Data visualization
# Boxplot of total gestures by group
ggplot(pretend_data_H10, aes(x = GroupStatus, y = total_gestures, fill = GroupStatus)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Total Gestures by Group",
       x = "Group Status",
       y = "Total Gesture Count")

# Bar chart of representational gesture presence by group
ggplot(pretend_data_H10, aes(x = GroupStatus, fill = factor(rep_gesture_present))) +
  geom_bar(position = "dodge") +
  scale_fill_discrete(name = "Rep. Gesture", labels = c("Absent", "Present")) +
  theme_minimal() +
  labs(title = "Representational Gesture Presence by Group",
       x = "Group Status",
       y = "Count")
```
# 3. Merge with narrative outcome data
```{r}
narrative_data <- read.csv(here("PN_Datasets", "Narrative", "CSVs_of_Combined_Data_N", "Narrative_Results_Final.csv"))

# Rename participant column
narrative_data <- narrative_data %>%
  rename(participant_id = ChildID)

write.csv(narrative_data, here("PN_DataAnalysis", "PP_Narrative_Analysis", "narrative_data.csv"))

# MERGE IT
completely_merged_data_H10 <- left_join(pretend_data_H10, narrative_data, by = "participant_id") %>%
  mutate(GroupStatus = if_else(GroupStatus == "BI", "PL", GroupStatus))

write_csv(completely_merged_data_H10, here("PN_DataAnalysis", "PP_Narrative_Analysis", "completely_merged_data_H10.csv"))
```

# 4. Run regression models
```{r}
# Model 1: Total gestures as predictor
model1 <- lm(AvgHighScore ~ GroupStatus + total_gestures + total_pretend_play + mlu + GroupStatus:total_gestures, 
             data = completely_merged_data_H10)

# Model 2: Representational gesture presence as predictor
model2 <- lm(AvgHighScore ~ GroupStatus + rep_gesture_present + total_pretend_play + mlu + GroupStatus:rep_gesture_present, 
             data = completely_merged_data_H10)

summary(model1)
summary(model2)

# Calculate effect sizes
eta_squared(model1)
eta_squared(model2)
```

# 5b. USE WHEN HAVE ALL DATA FROM 3 TIMEPOINTS..... mixed effects models 
```{r}
#library(lme4)
#library(lmerTest)

#mixed_model1 <- lmer(AvgHighScore ~ GroupStatus + total_gestures + total_pretend_play + mlu + 
                    # GroupStatus:total_gestures + (1|participant_id), 
                    # data = completely_merged_data)

#mixed_model2 <- lmer(AvgHighScore ~ GroupStatus + rep_gesture_present + total_pretend_play + mlu + 
                     #GroupStatus:rep_gesture_present + (1|participant_id), 
                     #data = completely_merged_data)

#summary(mixed_model1)
#summary(mixed_model2)
```


```{r}
# Interaction Plot: Group × Representational Gesture Presence
#| label: fig-interaction
#| fig-cap: "Predicted narrative scores by group and representational gesture use. Error bars represent 95% confidence intervals."
#| fig-width: 6
#| fig-height: 4
#| 
options(warn = 2)


# Create a data frame for prediction (numeric variables for model compatibility)
newdata <- expand.grid(
  GroupStatus = c("PL", "TD"),
  rep_gesture_present = c(0, 1),
  total_pretend_play = mean(completely_merged_data_H10$total_pretend_play, na.rm = TRUE),
  mlu = mean(completely_merged_data_H10$mlu, na.rm = TRUE)
)

# Get predictions and standard errors from the model
newdata$predicted <- predict(model2, newdata = newdata)
pred_se <- predict(model2, newdata = newdata, se.fit = TRUE)$se.fit
newdata$lower <- newdata$predicted - 1.96 * pred_se
newdata$upper <- newdata$predicted + 1.96 * pred_se

# Relabel variables as factors for plotting clarity
newdata$GroupStatus <- factor(newdata$GroupStatus, 
                              levels = c("PL", "TD"),
                              labels = c("Peri- and Postnatal Lesions", "Typically Developing"))
newdata$rep_gesture_present <- factor(newdata$rep_gesture_present,
                                      levels = c(0, 1),
                                      labels = c("No", "Yes"))

# Create the interaction plot
ggplot(newdata, aes(x = rep_gesture_present, y = predicted, 
                    group = GroupStatus, color = GroupStatus, 
                    linetype = GroupStatus, shape = GroupStatus)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  labs(x = "Representational Gestures Present",
       y = "Predicted Narrative Score") +
  scale_color_grey(start = 0.2, end = 0.8) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_shape_manual(values = c(16, 17)) +
  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black")
  )
```



# Scatter Plot: Raw Data with Regression Lines
#| label: fig-scatter
#| fig-cap: "Relationship between representational gesture presence and narrative scores by group. Shaded areas represent 95% confidence intervals around regression lines."
#| fig-width: 7
#| fig-height: 4

ggplot(completely_merged_data_H10, aes(x = as.factor(rep_gesture_present), y = AvgHighScore, 
                                 fill = GroupStatus, shape = GroupStatus)) +
  geom_jitter(width = 0.1, height = 0, size = 3, alpha = 0.7) +
  geom_smooth(aes(group = GroupStatus, linetype = GroupStatus, color = GroupStatus),
              method = "lm", se = TRUE, formula = y ~ x) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_grey(start = 0.5, end = 0.8) +
  scale_color_grey(start = 0.2, end = 0.6) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  labs(x = "Representational Gestures Present",
       y = "Narrative Score") +
  facet_wrap(~ GroupStatus, scales = "free_x") +
  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 12)
  )
```

```
# Total Gestures Scatter Plot with facet_grid
#| label: fig-total-gestures-grid
#| fig-cap: "Relationship between total gesture count and narrative scores, faceted by group status and the presence of representational gesture. The top row shows children who used representational gestures during pretend play, while the bottom row shows those who did not."
#| fig-width: 8
#| fig-height: 6

# First, create a factor for rep_gesture_present for better labeling
completely_merged_data$rep_gesture_factor <- factor(
  completely_merged_data$rep_gesture_present,
  levels = c(0, 1),
  labels = c("No Representational", "Representational")
)

ggplot(completely_merged_data, aes(x = total_gestures, y = AvgHighScore, 
                                 color = GroupStatus, shape = GroupStatus)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(aes(linetype = GroupStatus), method = "lm", se = TRUE) +
  labs(x = "Total Number of Gestures",
       y = "Narrative Score") +
  scale_color_grey(start = 0.2, end = 0.8) +
  scale_shape_manual(values = c(16, 17)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  # Add facet_grid with rep_gesture_factor in rows and GroupStatus in columns
  facet_grid(rep_gesture_factor ~ GroupStatus, scales = "free_x") +
  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 11)
  )
```