---
title: "PretendPlay_Gesture_Analysis"
author: "Kristen Johnson"
date: "2025-04-02"
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE}
# Install required packages if you don't have them
# Remove the comment symbols (#) to run these lines
# install.packages("tidyverse")
 install.packages("car")
 install.packages("effectsize")
 install.packages("rstatix")

# Load necessary packages
library(tidyverse)
library(car)       # For Levene's test and Type III SS
library(effectsize) # For effect sizes
library(rstatix)    # For convenient statistical functions
```

# My data is in a dataframe called 'merged_[timepoint]' with columns:
## - GroupStatus: Factor with levels "TD" and "PL"
## - gesture_all: Count of all gestures
## - gesture_rep: Count of representational gestures
## - gesture_icon: Count of iconic gestures
## - C-wpu: Language ability measure
## - c_pret: Frequency of pretend play instances

# 1. Conduct ANOVAs for each gesture type & summarize results
```{r}
# read in data frame that is just gesture during instance of pretend play
pretend_data_H8 <- read.csv("PN_Datasets/Pretend_Play/CSVs_of_Combined_Data_PP/pretend_data_H8.csv")

# Aggregate by child
child_summary <- pretend_data_H8 %>%
  group_by(participant_id, GroupStatus) %>% # Assuming you have a child_id variable
  summarize(
    total_gestures = sum(gesture_all),
    rep_gestures = sum(gesture_rep),
    icon_gestures = sum(gesture_icon),
    prop_rep = sum(gesture_rep) / sum(gesture_all),
    prop_icon = sum(gesture_icon) / sum(gesture_all)
  )

# For total gestures
total_anova <- aov(total_gestures ~ GroupStatus, data = child_summary)
cat("\nTotal gestures ANOVA:\n")
print(summary(total_anova))

# For representational gestures
rep_anova <- aov(rep_gestures ~ GroupStatus, data = child_summary)
cat("\nRepresentational gestures ANOVA:\n")
print(summary(rep_anova))
 
# For iconic gestures
icon_anova <- aov(icon_gestures ~ GroupStatus, data = child_summary)
cat("\nIconic gestures ANOVA:\n")
print(summary(icon_anova))
 
# For proportion of representational gestures to total gestures
prop_rep_anova <- aov(prop_rep ~ GroupStatus, data = child_summary)
cat("\nProportion of representational gestures ANOVA:\n")
print(summary(prop_rep_anova))
 
#for proportion of iconic gestures to total gestures
prop_icon_anova <- aov(prop_icon ~ GroupStatus, data = child_summary)
cat("\nProportion of iconic gestures ANOVA:\n")
print(summary(prop_icon_anova))

```

# 2. Calculate effect sizes
```{r}
if(require(effectsize)) {
  cat("\nEffect sizes:\n")
  cat("Total gestures: ")
  print(eta_squared(total_anova))
  cat("Representational gestures: ")
  print(eta_squared(rep_anova))
  cat("Iconic gestures: ")
  print(eta_squared(icon_anova))
  cat("Proportion representational: ")
  print(eta_squared(prop_rep_anova))
  cat("Proportion iconic: ")
  print(eta_squared(prop_icon_anova))
} else {
  # Manual calculation if package not available
  cat("\nEffect sizes calculated manually:\n")
  # Formula for eta-squared: SS_between / SS_total
  summary_total <- summary(total_anova)
  eta_sq_total <- summary_total[[1]]["GroupStatus", "Sum Sq"] / 
                 sum(summary_total[[1]][, "Sum Sq"])
  cat("Total gestures eta-squared: ", eta_sq_total, "\n")
  # Repeat for other ANOVAs
}
```

# 3. Descriptive statistics by group
```{r}
group_stats <- child_summary %>%
  group_by(GroupStatus) %>%
  summarize(
    n = n(),
    total_mean = mean(total_gestures, na.rm = TRUE),
    total_sd = sd(total_gestures, na.rm = TRUE),
    rep_mean = mean(rep_gestures, na.rm = TRUE),
    rep_sd = sd(rep_gestures, na.rm = TRUE),
    icon_mean = mean(icon_gestures, na.rm = TRUE),
    icon_sd = sd(icon_gestures, na.rm = TRUE),
    prop_rep_mean = mean(prop_rep, na.rm = TRUE),
    prop_rep_sd = sd(prop_rep, na.rm = TRUE),
    prop_icon_mean = mean(prop_icon, na.rm = TRUE),
    prop_icon_sd = sd(prop_icon, na.rm = TRUE)
  )

print("Descriptive statistics by group:")
print(group_stats)
```


# 4. Visualization of results
```{r}
if(require(tidyverse)) {
  # Raw counts visualization
  counts_long <- child_summary %>%
    select(GroupStatus, total_gestures, rep_gestures, icon_gestures) %>%
    pivot_longer(cols = c(total_gestures, rep_gestures, icon_gestures),
                names_to = "gesture_type", 
                values_to = "count")
  
  p1 <- ggplot(counts_long, aes(x = gesture_type, y = count, fill = GroupStatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.9)) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.9)) +
    labs(title = "Mean Gesture Counts During Pretend Play by Group",
         x = "Gesture Type",
         y = "Mean Count",
         fill = "Group Status") +
    theme_minimal() +
    scale_x_discrete(labels = c("total_gestures" = "Total",
                               "rep_gestures" = "Representational",
                               "icon_gestures" = "Iconic"))
  print(p1)
  
  # Proportions visualization
  props_long <- child_summary %>%
    select(GroupStatus, prop_rep, prop_icon) %>%
    pivot_longer(cols = c(prop_rep, prop_icon),
                names_to = "proportion_type", 
                values_to = "proportion")
  
  p2 <- ggplot(props_long, aes(x = proportion_type, y = proportion, fill = GroupStatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.9)) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.9)) +
    labs(title = "Proportion of Gesture Types During Pretend Play by Group",
         x = "Proportion Type",
         y = "Mean Proportion",
         fill = "Group Status") +
    theme_minimal() +
    scale_x_discrete(labels = c("prop_rep" = "Representational/Total",
                               "prop_icon" = "Iconic/Total")) +
    scale_y_continuous(labels = scales::percent)
  print(p2)
}
```

# 5. Statistical tests to directly compare TD and BI groups
```{r}
# t-tests for each variable (alternative to ANOVA with only two groups)
t_total <- t.test(total_gestures ~ GroupStatus, data = child_summary)
t_rep <- t.test(rep_gestures ~ GroupStatus, data = child_summary)
t_icon <- t.test(icon_gestures ~ GroupStatus, data = child_summary)
t_prop_rep <- t.test(prop_rep ~ GroupStatus, data = child_summary)
t_prop_icon <- t.test(prop_icon ~ GroupStatus, data = child_summary)

cat("\nt-test results (direct comparison between groups):\n")
cat("\nTotal gestures:\n")
print(t_total)
cat("\nRepresentational gestures:\n")
print(t_rep)
cat("\nIconic gestures:\n")
print(t_icon)
cat("\nProportion representational:\n")
print(t_prop_rep)
cat("\nProportion iconic:\n")
print(t_prop_icon)
```

# 6. Calculate Cohen's d effect sizes for t-tests with error handling
```{r}

cat("\nCohen's d effect sizes:\n")

# Function to safely calculate Cohen's d
safe_cohens_d <- function(formula, data) {
  tryCatch({
    res <- cohens_d(formula, data = data)
    return(res)
  }, error = function(e) {
    # Extract variable name from formula
    var_name <- as.character(formula)[2]
    # Get means by group
    means <- aggregate(formula, data = data, FUN = mean, na.rm = TRUE)
    sds <- aggregate(formula, data = data, FUN = sd, na.rm = TRUE)
    
    cat("Error calculating Cohen's d for", var_name, ":\n")
    cat("Group means:", toString(means), "\n")
    cat("Group SDs:", toString(sds), "\n")
    cat("Error message:", e$message, "\n")
    return(NULL)
  })
}

# Apply the safe function to each variable
cat("\nTotal gestures:\n")
print(safe_cohens_d(total_gestures ~ GroupStatus, data = child_summary))

cat("\nRepresentational gestures:\n")
print(safe_cohens_d(rep_gestures ~ GroupStatus, data = child_summary))

cat("\nIconic gestures:\n")
print(safe_cohens_d(icon_gestures ~ GroupStatus, data = child_summary))

cat("\nProportion representational:\n")
print(safe_cohens_d(prop_rep ~ GroupStatus, data = child_summary))

cat("\nProportion iconic:\n")
print(safe_cohens_d(prop_icon ~ GroupStatus, data = child_summary))
```

# Make APA-formatted data visualizations
```{r}
if(require(tidyverse)) {
  # Set APA theme
  apa_theme <- theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(colour = "black", fill = NA),
      axis.text = element_text(colour = "black", size = 12),
      axis.title = element_text(size = 12, face = "bold"),
      legend.title = element_text(size = 12, face = "bold"),
      legend.text = element_text(size = 12),
      legend.position = "bottom",
      legend.background = element_rect(fill = "white", colour = "black"),
      plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
      text = element_text(family = "Times New Roman")
    )
  
  # Raw counts visualization
  counts_long <- child_summary %>%
    select(GroupStatus, total_gestures, rep_gestures, icon_gestures) %>%
    pivot_longer(cols = c(total_gestures, rep_gestures, icon_gestures),
                names_to = "gesture_type", 
                values_to = "count")
  
  # Renaming GroupStatus labels to be more readable
  counts_long$GroupStatus <- factor(counts_long$GroupStatus, 
                                  levels = c("BI", "TD"),
                                  labels = c("Brain Injury", "Typically Developing"))
  
  p1 <- ggplot(counts_long, aes(x = gesture_type, y = count, fill = GroupStatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.8), 
                alpha = 0.8, color = "black", size = 0.2) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.8)) +
    labs(title = "Figure 1",
         subtitle = "Mean Gesture Counts During Pretend Play by Group",
         x = "Gesture Type",
         y = "Mean Count",
         fill = "Group") +
    scale_x_discrete(labels = c("total_gestures" = "Total",
                               "rep_gestures" = "Representational",
                               "icon_gestures" = "Iconic")) +
    scale_fill_grey(start = 0.4, end = 0.8) +
    apa_theme +
    theme(plot.subtitle = element_text(size = 12, hjust = 0.5))
  
  print(p1)
  
  # Save high-resolution figure
  ggsave("Figure_1_Gesture_Counts.tiff", p1, width = 7, height = 5, dpi = 300)
  
  # Proportions visualization
  props_long <- child_summary %>%
    select(GroupStatus, prop_rep, prop_icon) %>%
    pivot_longer(cols = c(prop_rep, prop_icon),
                names_to = "proportion_type", 
                values_to = "proportion")
  
  # Renaming GroupStatus labels
  props_long$GroupStatus <- factor(props_long$GroupStatus, 
                                  levels = c("BI", "TD"),
                                  labels = c("Brain Injury", "Typically Developing"))
  
  p2 <- ggplot(props_long, aes(x = proportion_type, y = proportion, fill = GroupStatus)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(0.8), 
                alpha = 0.8, color = "black", size = 0.2) +
    stat_summary(fun.data = function(x) {
      data.frame(y = mean(x, na.rm = TRUE),
                ymin = mean(x, na.rm = TRUE) - sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))),
                ymax = mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)/sqrt(sum(!is.na(x))))
    }, geom = "errorbar", width = 0.2, position = position_dodge(0.8)) +
    labs(title = "Figure 2", 
         subtitle = "Proportion of Gesture Types During Pretend Play by Group",
         x = "Proportion Type",
         y = "Mean Proportion",
         fill = "Group") +
    scale_x_discrete(labels = c("prop_rep" = "Representational/Total",
                               "prop_icon" = "Iconic/Total")) +
    scale_y_continuous(labels = scales::percent, 
                      breaks = seq(0, 0.25, 0.05)) +
    scale_fill_grey(start = 0.4, end = 0.8) +
    apa_theme +
    theme(plot.subtitle = element_text(size = 12, hjust = 0.5))
  
  print(p2)
  
  # Save high-resolution figure
  ggsave("Figure_2_Gesture_Proportions.tiff", p2, width = 7, height = 5, dpi = 300)
}
```

